---
title: "Prepare data - neurogenesis"
output: html_notebook
---

# Download Llorens-Bobadilla 2015 dataset

The data is downloaded on on 03/01/2019.

Load data from [Dulken et al. 2017](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC5269583/)

This code is obtained from the [GitHub code for Figure 6B in Dulken et al 2017](https://github.com/bdulken/SVZ_NSC_Dulken_2/blob/master/Figure6_FigureS6/Llorens%20Projections%20_%20Figure%206B%2C%20S6A%2C%20S6B.R)

```{r}
Llorens_allcounts<-read.table(url("https://github.com/bdulken/SVZ_NSC_Dulken_2/raw/master/Files/Llorens_counts_allgenes.txt"))

allcounts_allcells<-Llorens_allcounts

#Remove neuroblasts from Llorenss data
#allcounts_allcells_notaps<-allcounts_allcells[!grepl("tap",colnames(allcounts_allcells))]
allcounts_allcells_noblasts<-allcounts_allcells[!grepl("PSA",colnames(allcounts_allcells))]

#Remove oligodendrocytes from Llorenss data
oligos<-as.vector(read.table(url("https://raw.githubusercontent.com/bdulken/SVZ_NSC_Dulken_2/master/Files/comp_oligos.txt"))[,1])
allcounts_allcells_noblasts_nooligo<-allcounts_allcells_noblasts[,-na.omit(match(oligos,colnames(allcounts_allcells_noblasts)))]
allcounts_allcells_noblasts_nooligo_noERCC<-allcounts_allcells_noblasts_nooligo[!grepl("ERCC-",rownames(allcounts_allcells_noblasts_nooligo)),]

llorens <- allcounts_allcells_noblasts_nooligo_noERCC

llorens_cell_type <-  factor(
  unlist(
    lapply(X = colnames(llorens),
           FUN = function(x){substr(x,1,1)})
    )
  )

llorens_cell_type <- plyr::mapvalues(x = llorens_cell_type,
                                     from = c("N","t"),
                                     to = c("NSC","NPC"))

llorens <- SingleCellExperiment(assay = list(counts = as.matrix(llorens)))
llorens$cell_type1 <- llorens_cell_type

saveRDS(llorens, file = "sce_rds_files/llorens.Rds")
```





# Download Dulken 2017 dataset

The data is downloaded on on 03/04/2019.



This code is obtained from the [GitHub code for Figure 1 in Dulken et al 2017](https://github.com/bdulken/SVZ_NSC_Dulken_2/blob/master/Figure1/PCA%20-%20Outliers%20Removed%20_%20Figure%201C.R). The difference between our code and theirs is that we don't filter out cell-cycle genes.


```{r}
#Loading all high quality cells and filtering for lowly expressed genes
spec_pops<-read.table(url("https://raw.githubusercontent.com/bdulken/SVZ_NSC_Dulken_2/master/Files/AllCounts_specPops_read_gene_ERCC_filt_FINAL.txt"))

allcounts_allcells<-spec_pops

#Removing Oligodendrocytes and Outliers
oligos<-as.vector(read.table(url("https://raw.githubusercontent.com/bdulken/SVZ_NSC_Dulken_2/master/Files/STAR_oligos_updated_09232015.txt"))[,1])
allcounts_allcells_nooligo<-allcounts_allcells[,-na.omit(match(oligos,colnames(allcounts_allcells)))]

#Filtering for expressed by 5 cells at 10 counts
greaterthan0<-allcounts_allcells_nooligo>10
greaterthan0sum<-rowSums(greaterthan0)
allcounts_allcells_nooligo_genefilt<-allcounts_allcells_nooligo[greaterthan0sum>=5,]
dulken <- allcounts_allcells_nooligo_genefilt

# Overview of the Dulken dataset
dulken_cell_type <- 
  factor(
    unlist(
      lapply(X=colnames(dulken),
             FUN=function(x){strsplit(x,split = "_", fixed=T)[[1]][1]})
      )
    )

# TAP stands for transient amplifying cells, which is equivalent to neural progenitor cells here
dulken_cell_type <- plyr::mapvalues(x = dulken_cell_type,
                              from = c("aNSC", "Ast", "NPC", "qNSC", "NSC", "TAP"),
                              to = c("NSC","Ast","NPC","NSC","NSC","NPC"))

library(SingleCellExperiment)

dulken <- SingleCellExperiment(assay = list(counts = as.matrix(dulken)))
dulken$cell_type1 <- dulken_cell_type
saveRDS(dulken, file = "sce_rds_files/dulken.Rds")
```




# Download Dulken 2017 consensus genes

```{r}
library(openxlsx)

temp <- tempfile()
download.file("https://www.cell.com/cms/10.1016/j.celrep.2016.12.060/attachment/ba11d4f6-39dc-4029-a0a3-73581739daf1/mmc9.xlsx",temp)

sheet_names <- openxlsx::getSheetNames(temp)

lapply(sheet_names,
       function(sheet_name){
         read.xlsx(temp, sheet = sheet_name,startRow = 3, rowNames = T)
       }) -> 
  table_S8

names(table_S8) <- sheet_names
saveRDS(table_S8, file = "other_rds_files/table_S8.Rds")

unlink(temp)
```


We download the gene ranking produced by a supervised feature selection method as in [Dulken et al. 2017](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC5269583/)

Note:

- The table `consensus_genes` has 101 columns, the first column is metadata
- The remainder columns are gene rankings computed by Dulken et al.
- Each column is the top 100 genes produced by one computed model.
- A gene that shows up in multiple columns means that there is a _consensus_ across multiple models that the gene should be ranked in the top 100, hence the name.

```{r, eval = F}
library("openxlsx")
library(dplyr)
consensus_genes <- 
  read.xlsx("https://www.ncbi.nlm.nih.gov/pmc/articles/PMC5269583/bin/NIHMS839312-supplement-7.xlsx")
consensus_genes <- consensus_genes[,-c(1)] # remove the first column, which is metadata


all_consensus_genes <- Reduce(union, consensus_genes)

all_consensus_genes %>%
  sapply(function(g){
    # count up how many models does g show up in
    consensus_genes %>%
      sapply(function(model) {
        g %in% model
      }) %>%
      sum
  }) %>% 
  sort(decreasing = T) -> # top of the list are genes that show up the most and vice versa
  consensus_genes
```

```{r}
dulken <- readRDS(file = "sce_rds_files/dulken.Rds")
llorens <- readRDS(file = "sce_rds_files/llorens.Rds")
shared_genes <- intersect(rownames(dulken), rownames(llorens))
consensus_genes <- consensus_genes[names(consensus_genes) %in% shared_genes]

saveRDS(consensus_genes, file = "other_rds_files/consensus_genes.Rds")
```